---
title: "02_Bootstrapping"
author: "Akriti"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    toc: true
    number_section: true
    toc_float: 
      collapsed: false
      smooth_scroll: true
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../results/") })
---

## Barcode abundance analysis

#Step 1. Barcode quantification

We found result of sequencing as a fastq files in folders (as shown on hpc as HGWKVAFX5). We have two type of reads HGWKVAFX5_01 that contains Read1 and HGWKVAFX5_02 that contains Read_2. 

#Some details about the sequences

1. This sequencing result is a PCR_product sequencing result and we know we expect some 50 bases in the sequence (could be seen in the file BARSEQ_YETI_Library in SnapGene). 
Only Read_1 is important to us because we expect it to contain Barcodes. 
2. The barcode sequence are 12bp long. 
3. We performed quantification (counting the number of barcodes in each experiment/timepoints) on hpc, we merged the files and extracted the final count file which is loaded in this rmd. 

#Loading the required library

```{r, include = F, Echo = F}
#BiocManager::install("clusterProfiler")

#BiocManager::install("enrichplot")

library(stringr)
library(dplyr)
library(tibble)
library(purrr)
library(ggplot2)
library(tidyverse)
library(Rsamtools)
library(GenomicAlignments)
library(BiocParallel)
library(DESeq2)
library(pheatmap)
library(ComplexHeatmap)
library(clusterProfiler)
library(enrichplot)
library(tidyverse)
```

##Load the count file

From this table we can see the first three columns of the count files have information for **Barcode**, its associated **ORF** and **Gene**. The other columns are different samples collected at different time points. 
```{r}
merged_barcode_counts <- read.table("../data/Barcode_counts", quote="\"", comment.char="")

str(merged_barcode_counts)
head(merged_barcode_counts)


```
##Editing the count table. 

This step includes renaming the column, replacing NA's to 0, and removal of the last column that contains undetermined sequences. 
```{r}
#Rename columns of merged barcode counts
colnames(merged_barcode_counts) <- c("Barcode", 
                                     "ORF", 
                                     "Gene", 
                                     "H_0", 
                                     "YPD_12H.1", 
                                     "YPD_12H.2", 
                                     "E2_12H.1", 
                                     "E2_12H.2", 
                                     "E2_12H_Flu.1", 
                                     "E2_12H_Flu.2", 
                                     "E2_12H_Flu.3",
                                     "YPD_24H.1", 
                                     "YPD_24H.2", 
                                     "E2_24H_Flu.2", 
                                     "E2_24H.2", 
                                     "E2_24H_Flu.1", 
                                     "E2_24H.1", 
                                     "YPD_48H.1", 
                                     "YPD_48H.2",
                                     "E2_48H.1", 
                                     "E2_48H.2",
                                     "E2_48H_Flu.1", 
                                     "E2_48H_Flu.2", 
                                     "YPD_72H.1", 
                                     "YPD_72H.2", 
                                     "E2_72H.1", 
                                     "E2_72H.2", 
                                     "E2_72H_Flu.1", 
                                     "E2_72H_Flu.2", 
                                     "undetermined")

#rearrange the columns - While performing experiment, the 24h sample for 100Est, 0flu and 100Est, 32 flu was interchanged because of labelling error, since, my values were also indicating this error confirmed by very high count in the E2 col which I expect in E2-flu column. I replaced that dataset. 
merged_barcode_counts <- merged_barcode_counts %>%  dplyr ::select(Barcode, ORF,
                                     Gene,
                                     "H_0",
                                     "YPD_12H.1", 
                                     "YPD_12H.2", 
                                     "E2_12H.1", 
                                     "E2_12H.2", 
                                     "E2_12H_Flu.1", 
                                     "E2_12H_Flu.2", 
                                     "E2_12H_Flu.3",
                                     "YPD_24H.1", 
                                     "YPD_24H.2", 
                                     "E2_24H.1", 
                                     "E2_24H.2", 
                                     "E2_24H_Flu.1", 
                                     "E2_24H_Flu.2",
                                     "YPD_48H.1", 
                                     "YPD_48H.2",
                                     "E2_48H.1", 
                                     "E2_48H.2",
                                     "E2_48H_Flu.1", 
                                     "E2_48H_Flu.2", 
                                     "YPD_72H.1", 
                                     "YPD_72H.2", 
                                     "E2_72H.1", 
                                     "E2_72H.2", 
                                     "E2_72H_Flu.1", 
                                     "E2_72H_Flu.2")

                                    
#Replace NA to 0 from column 4th to last
merged_barcode_counts[, 4:29][is.na(merged_barcode_counts[, 4:29])] <- 0
str(merged_barcode_counts)

#filtering the rows with count values greater than 10
#merged_barcode_counts <- merged_barcode_counts[merged_barcode_counts$E2_72H.1 > 1 & merged_barcode_counts$E2_72H_Flu.2 >10,]


```

##Create a metadata table

This step includes analysis of the count data matrix and sticking the extracted information to a metadata table. For each experiment, we specify the conditions, timepoints, replicates, the number of barcode count detected, the mean and sd of the barcode counts detected, and total read counts. 
```{r}
#Column 1 of metadata is  
experiments = colnames(merged_barcode_counts)[4:ncol(merged_barcode_counts)]

#Column 2 of metadata is 
timepoints = factor(c("0H", rep("12H", times=7), rep("24H", times =6), rep("48h", times =6), rep("72h", times=6)))

#Column 3 of metadata is 
conditions <- factor(c(rep("YPD", times=3), 
                       rep("E2", times =2), 
                       rep("E2-flu", times =3), 
                       rep("YPD", times=2), 
                       rep("E2", times =2),
                       rep("E2-flu", times =2), 
                       rep("YPD", times=2), 
                       rep("E2", times =2),
                       rep("E2-flu", times =2), 
                       rep("YPD", times=2), 
                       rep("E2", times =2), 
                       rep("E2-flu", times =2)))
#Here, YPD : No estradiol or fluconazole; E2 : 100nM estradiol and 0 ug/ml fluconazole; E2-flu: 100nM estradiol and 32 ug/ml fluconazole

#column 4 of metadata is
replicates <- factor(c(1,1,2,1,2,1,2,3,1,2,1,2,1,2,1,2,1,2,1,2,1,2,1,2,1,2))

#column 5 of metadata is the number of barcodes/genotypes detected in each experiments. This loop will count the number of times we detected any barcode, if suppose for a row we have a value greater than 0, it will take that column as 1, i.e., the read detected. It won't count the number

barcode_counts <- numeric(ncol(merged_barcode_counts)-3) #create an empty vector, exclude the first three character columns and verify the result using length()
for (i in 4:ncol(merged_barcode_counts)) {
  barcode_counts[i-3] <- sum(merged_barcode_counts[,i]>0)
} #loop through column 4 or 4th indices, which is why we provide i-3 everywhere in the for loop


#column 6 of metadata is the means of barcode counts per experiment
barcode_means <- numeric(ncol(merged_barcode_counts)-3)
for(i in 4:ncol(merged_barcode_counts)){
  barcode_means[i-3] <- mean(merged_barcode_counts[,i])
}

#column 7 of metadata is the sd of barcode counts per experiment
barcode_sd <- numeric(ncol(merged_barcode_counts)-3)
for(i in 4:ncol(merged_barcode_counts)){
  barcode_sd[i-3] <- sd(merged_barcode_counts[,i])
}

#column 8 of metadata is the total number of sum of counts of each sample, i.e., the total number of read counts
total_barcodes_per_experiment <- numeric(ncol(merged_barcode_counts)-3)
for(i in 4:ncol(merged_barcode_counts)){
  total_barcodes_per_experiment[i-3] <- sum(merged_barcode_counts[,i])
}

#Metadata
Metadata <- data.frame(Experiments = experiments, Timepoints = timepoints, Conditions = conditions, Replicates= replicates, Number_of_barcodes_detected = barcode_counts, means_of_barcodes_detected = barcode_means, Standard_deviation_of_barcodes_detected = barcode_sd, Read_counts = total_barcodes_per_experiment)

head(Metadata)
```

##The barcodes detected across the experiments

From this histogram, we can see that nearly 5000 barcodes are detected across every experiments performed at different timepoints. Since, our genotype repository is nearly 5000, this indicates that sequencing was successful and most barcodes were detected. Moreover, the approximate uniformity across different experimental conditions also show that the experiment and sequencing was efficiently executed. 

```{r}
ggplot(data = Metadata, aes(x = Experiments, y = Number_of_barcodes_detected)) +
  geom_bar(stat = "identity", fill = "orange") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Timepoint", y = "Detected Barcodes") +
  ggtitle("Histogram of Detected Barcodes by Timepoint")
```

##The read counts detected in each time points.

This plot shows the sequencing depth for each experiment. As we can see many experiments/time points have total 5M reads but we also observe range from 3M reads to 6M reads in different time points. Two of the very different read count is at 12h when treated with estradiol and fluconazole and 24hr estradiol and fluconazole treated one. 
```{r}
ggplot(data = Metadata, aes(x = Experiments, y = Read_counts)) +
  geom_bar(stat = "identity", fill = "orange") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Timepoint", y = "Read Counts") +
  ggtitle("Histogram of Detected Barcodes by Timepoint")
```
##Extract ERG11 count data fro all experiments/timepoints

ERG11 is a lanosterol demethylase, it demethylates lanosterol to cholestrol in ergosterol biosynthetic pathway. Azoles block the demethylation of lanosterol and inturn inhibit the ergosterol synthesis. Mutations in ERG11 is associated with fluconazole resistance. It is reported that upregulation of ERG11 is associated with fluconazole resitance, since, it can effectively decrease intracellular drug accumulation(Paul 2019; https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6910709/). 

In this analysis, ERG11 is our control and we expect an increase in ERG11 count as we treat the cells with estradiol and fluconazole. i.e., we expect that if the experiments was successful the overexpression of the ERG11 in the presence of E2 and Flu, will results in increase in abundance of the ERG11 genotype/strain that confers the increase in tolerance to fluconazole. Bascially, the increase in ERG11 expression desensitizes the cells to fluconazole. 

In this analysis we expect and report to see an increase in abundance of ERG11 in the presence of both E2-flu when compared to only E2 and no E2 and Flu.

```{r}
erg11_counts <- merged_barcode_counts %>% 
  dplyr::filter(Gene == "ERG11") %>% 
  dplyr::select(4:ncol(merged_barcode_counts))

ERG11_counts <- data.frame(t(erg11_counts))
colnames(ERG11_counts) <- "ERG11_barcode_count"
ERG11_counts <- rownames_to_column(ERG11_counts, var = "Experiments")

#merge ERG11_counts to the metadata
Metadata <- merge(Metadata, ERG11_counts, by = "Experiments")

ggplot(Metadata, aes(x = Timepoints, y = ERG11_barcode_count, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "ERG11 counts variation across experiments and conditions",
       x = "Experiment", y = "ERG11_counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_log10()
```
##Part 2

##Objective: To report which genotype increased in abundance in the presence, of fluconazole out of 5000 genotypes assessed. 

Step 1. We assessed the **genotype abundance** across **conditions** YPD, E2 and E2-YPD at each time point 12h, 24h, 48h, 72h. To do this we used the package deseq2 which will internally perform normalization and we can get the result for logfoldchange of different barcodes. We studied genotype in two contrasts such as *Flu v/s E2* and *YPD v/s E2*. The log2foldchange in the contrast 1, *Flu v/s E2* will show deseq results for the genotype aboundance in the presence of estradiol-fluconazole treated cells when compared to only estradiol treated cells. Similarly, in the contrast 2, *YPD v/s E2* will show deseq results for the genotype abundance in the YPD relative to estradiol. 

In the following code chunk remove the character columns from the merge_barcode_count dataframe because we need columns with same datatype for the deseq analysis. 
```{r}
#Add rownames as ORF
merge_col <- merged_barcode_counts[,-2]
row.names(merge_col) <- merged_barcode_counts[,2]
#remove the character columns 
all_time_points <- merge_col[,-c(1:3)]

#added by saaz 6/24/26
merge_col2=merge_col
merge_col2[,c(3:ncol(merge_col2))]=merge_col2[,c(3:ncol(merge_col2))]+1
```


##2. The following analysis reports the genotype abundance at 12h timepoint 

For deseq analysis, we need 12h count table, we need conditions table or coldata and design set as ~conditions. 

Using the following code chunk we obtained the log2fold change of genotpye in different conditions when compared to E2 as a reference. Although most highly abundant genes are not statistically significant. 
If we sum the padj value for the differentially expressed genes/log2foldchange in Flu/E2 contrast it is 0, therefore we chose to not include this data in the heatmap step or in the comparison of genotype abundance across timepoint.
```{r}
#1.extract count data with the 12h timepoint
merged_barcode_counts_12h <- merge_col2[,c(4:10)]
h_12 <- (merged_barcode_counts_12h)
#2. coldata
col1 <- names(merged_barcode_counts_12h)
condition = c("YPD", "YPD", "E2", "E2", "Flu", "Flu","Flu")
df <- data.frame(col1, condition)
rownames(df) <- df[,1]
coldata <- subset(df, select = -col1)
coldata
#factor the condition column
coldata$condition <- factor(coldata$condition)
levels(coldata$condition)

#3.Create a deseq object
dds12 <- DESeqDataSetFromMatrix(countData = h_12,
                              colData = coldata,
                              design = ~ condition)
dds12

#4.Run deseq
dds12 <- DESeq(dds12)
res <- results(dds12)
res

#5.Contrast is used to extract the tables with Flu/E2 and YPD/E2, where E2 is the reference to compare
resultsNames(dds12)
res1 <- results(dds12, contrast=c("condition", "Flu", "E2"))
res2 <- results(dds12, contrast=c("condition", "YPD", "E2"))


#6. find the stastically significant differentially expressed genes
sum(res1$pvalue < 0.05, na.rm = T)
sum(res1$padj < 0.05, na.rm = T)

#We did not find any significant fluconazole resistant genes at 12 hours
```

##3. Similarly, we repeat this analysis to report the genotype abundance at 24h timepoint

Using the following code chunk we obtained the log2fold change of genotpye in different conditions when compared to E2 as a reference. Although most highly abundant genes are not statistically significant. 
If we sum the padj value for the differentially expressed genes/log2foldchange in Flu/E2 contrast it is 0, therefore we chose to not include this data in the heatmap step or in the comparison of genotype abundance across timepoint
```{r}
#1.extract count data with the 24h timepoint
merged_barcode_counts_24h <- merge_col2[,c(11:16)]
h_24 <- (merged_barcode_counts_24h)
#2. coldata
col2 <- names(merged_barcode_counts_24h)
condition2 = c("YPD", "YPD", "E2", "E2", "Flu", "Flu")
df2 <- data.frame(col2, condition2)
rownames(df2) <- df2[,1]
coldata2 <- subset(df2, select = -col2)
#factor the condition column
coldata2$condition2 <- factor(coldata2$condition2)
levels(coldata2$condition2)

#3.Create a deseq object
dds24 <- DESeqDataSetFromMatrix(countData = h_24,
                              colData = coldata2,
                              design = ~ condition2)
dds24

#4.Run deseq
dds24 <- DESeq(dds24)
res24 <- results(dds24)
res24

#5. Contrast is used to extract the tables with Flu/E2 and YPD/E2, where E2 is the reference to compare
resultsNames(dds24)
res24.1 <- results(dds24, contrast=c("condition2", "E2", "Flu"))
res24 <- results(dds24, contrast=c("condition2", "E2", "YPD"))

#6. find the stastically significant differentially expressed genes
sum(res24.1$pvalue < 0.05, na.rm = T)
sum(res24.1$padj < 0.05, na.rm = T)

#We did not find any significant fluconazole resistant genes at 24 hours
```

##4. Report the genotype abundance at 48h timepoint

Using the following code chunk we obtained the log2fold change of genotpye in different conditions when compared to E2 as a reference. Many highly abundant genes are statistically significant. 

If we sum the padj value for the differentially expressed genes/log2foldchange for Flu/E2, it is 4042
```{r}
#1.extract count data with the 48h timepoint
h_48 <- merge_col2[,c(17:22)]

#2. coldata
condition3 = factor(c("YPD", "YPD", "E2", "E2", "Flu", "Flu"))
coldata3 <- data.frame(row.names = names(h_48), condition3)

#factor the condition column
levels(coldata3$condition3)

#3.Create a deseq object
dds48 <- DESeqDataSetFromMatrix(countData = h_48,
                              colData = coldata3,
                              design = ~ condition3)
dds48

#4.Run deseq
dds48 <- DESeq(dds48)
res48 <- results(dds48)
res48
#5. Contrast is used to extract the tables with Flu/E2 and YPD/E2, where E2 is the reference to compare
resultsNames(dds48)
res48.1 <- results(dds48, contrast=c("condition3", "Flu", "E2"))
res48.2 <- results(dds48, contrast=c("condition3", "YPD", "E2"))

#6. find the stastically significant differentially expressed genes
sum(res48.1$padj < 0.05, na.rm = T) #4042
sum(res48.2$padj < 0.05, na.rm = T) #1483

res48.1["YMR202W",]

plotMA(res48.1)

```

##5. Report the genotype abundance at 72h timepoint

Using the following code chunk we obtained the log2fold change of genotpye in different conditions when compared to E2 as a reference. Many highly abundant genes are statistically significant. 

If we sum the padj value for the differentially expressed genes/log2foldchange for Flu/E2, it is 1745
```{r}
#1.extract count data with the 72h timepoint
h_72  <- merge_col2[,c(23:28)]

h_72["YMR202W",]
#2. coldata
condition4 = factor(c("YPD", "YPD", "E2", "E2", "Flu", "Flu"))
coldata4 <- data.frame(row.names = names(h_72), condition4)

#factor the condition column
levels(coldata4$condition4)

#3.Create a deseq object
dds72 <- DESeqDataSetFromMatrix(countData = h_72,
                              colData = coldata4,
                              design = ~ condition4)
dds72

#4.Run deseq
dds72 <- DESeq(dds72)
res72 <- results(dds72)
res72
#5. Contrast is used to extract the tables with Flu/E2 and YPD/E2, where E2 is the reference to compare
resultsNames(dds72)
res72.1 <- results(dds72, contrast=c("condition4", "Flu", "E2"))
res72.2 <- results(dds72, contrast=c("condition4", "YPD", "E2"))

#6. find the stastically significant differentially expressed genes
sum(res72.1$padj < 0.05, na.rm = T) #1745
sum(res72.2$padj < 0.05, na.rm = T) #0

res72.1["YMR202W",]

res72["YMR202W",]
```

##6 Heatmap for fluconazole v/s Estradiol across different timepoints. 

The objective for this heat map is to find which genotypes when overexpressed in the presence of estradiol show increase in abundance in the presence of fluconazole which shows that they are selected for. Essentially which are the high abundance and low abundance genotype which are present in both 48h and 72h Estradiol-fluconazole treated cells relative to only estradiol treated cells.
Each panel in this heatmap tells the the genotype abundant at 72h or at 48h or at 0h when Estradiol-fluconazole treated cells are contrasted to only estradiol treated cells. 
Red represents increase in abundance. Inference: Overexpression of the red genotype promotes desensitization to fluconazole as a result it confers increased tolerance to fluconazole. 
Blue represents decrease in abundance. Inference: Overexpression of the blue genotype promotes fitness defect i.e., it sensitizes the cells to fluconazole


#Heatmap for statistically significant log2foldchange values
The following plot describes the logfold2change of genotypes across the three timepoints in fLu v/s E2 contrast. It only contains gene which are statistically significant i.e., padj < 0.05 & log2foldpvalue>1.5. 

```{r}
#select the statistically significant log2foldchange values. changed 6/24/26 by saaz.
res48.1_filtered_for_statistical_significance <- res48.1[!is.na(res48.1$padj) & res48.1$padj < 0.05 & abs(res48.1$log2FoldChange) > 1.5, ]

#res48.1_filtered_for_statistical_significance <- res48.1[is.na(res48.1$padj) | res48.1$padj < 0.05, ]

result_48.1 <- data.frame(row.names = rownames(res48.1_filtered_for_statistical_significance), log2FoldChange = res48.1_filtered_for_statistical_significance$log2FoldChange)


#select the statistically significant log2foldchange values, changed 6/24/26 by saaz.
res72.1_filtered_for_statistical_significance <- res72.1[!is.na(res72.1$padj) & res72.1$padj < 0.05 & abs(res72.1$log2FoldChange) > 1.5, ]
result_72.1 <- data.frame(row.names = rownames(res72.1_filtered_for_statistical_significance), log2FoldChange = res72.1_filtered_for_statistical_significance$log2FoldChange)

#To differentiate the table with 48h values with 72 hours value we assign a column with time information
result_48.1$time <- 48
result_72.1$time <- 72

#convert gene/ORF column to a new column gene
result_48.1$gene <- rownames(result_48.1)
rownames(result_48.1) <- NULL

result_72.1$gene <- rownames(result_72.1)
rownames(result_72.1) <- NULL


#merge the two dataframe using the gene column
Significant_genes_48_72 <- merge(result_48.1, result_72.1, by="gene", all = TRUE)
#replace the NA values with 0
Significant_genes_48_72[is.na(Significant_genes_48_72)] <- 0


#Create a heatmap
rownames(Significant_genes_48_72) <- Significant_genes_48_72$gene
heatmap_genes_48_72 <- Significant_genes_48_72[,-c(1,3,5)]


heatmap_genes_48_72 <- heatmap_genes_48_72 %>% 
  mutate(time_0h = 0)

colnames(heatmap_genes_48_72) <- c("time_48h", "time_72h", "time_0h")
#rearrange the columns
heatmap_genes_48_72 <-heatmap_genes_48_72 %>%  dplyr::select(time_0h,
                                                     time_48h,
                                                     time_72h)
                                    
#hierarchical clustering to cluster the logfold change, it makes the heatmap more meaningful
hclust <- hclust(dist(heatmap_genes_48_72), method = "complete")
#heatmap
Heatmap(heatmap_genes_48_72, cluster_rows = hclust, show_row_names = F)

#rearrange the columns based on the decreasing order of 72h because we are looking for genotype that increased in abundance in the presence of fluconazole treatment. 72h is the longest time we experimented. 
decreasing_order_of_significant_genes <- heatmap_genes_48_72[ order(-heatmap_genes_48_72$time_72h), ]

#add common gene name column to this table
gene_name <- merged_barcode_counts[,2:3]
rownames(gene_name) <-gene_name[,1]
decreasing_order_of_significant_genes$gene <- gene_name[row.names(decreasing_order_of_significant_genes), "Gene"]




#View the significant genes in decreasing order
View(decreasing_order_of_significant_genes)
```


```
Q) What are the top significant genes in this analysis?
Ans: Some of the top significant genes are:
1. Most abundant -- FLR1 -- Fluconazole resistance
2. ERG11
3. PDR5
4. YAP1
5. ERG13

```

###PART 3 -- Extract important genes and plot the change in raw count across the experimental time points

#plot the results for specific genes -- 

Extract the top hit gene counts from the raw data table, merge it to the metadata for the plotting purposes.
```{r}
#filter the top hits
Gene.counts <- filter(merged_barcode_counts[,-(1:2)], Gene %in% c("FLR1", "ERG11", "PDR5", "YAP1", "ERG13", "MGE1", "PDR3", "SOK1", "MDJ1", "MRS4", "SKS1", "RPL27A", "NUP57", "RTS2", "STP4", "MSN5", "VAS1", "UPC2", "AUS1", "FRE3", "HSP104"))
#transpose for ggplot
Gene_counts <- setNames(data.frame(t(Gene.counts[,-1])), Gene.counts[,1])
Gene_counts <- rownames_to_column(Gene_counts, var = "Experiments")
#merge with metadata
Metadata <- merge(Metadata, Gene_counts, by = "Experiments")
```

#Data trimming to remove anomaly that occured due to technical error
```{r}
#Step 1 remove the E2_12H_Flu with 
Metadata <- Metadata[-c(1:2), ]

Metadata <- Metadata[-4,]
```





#Visualization of the raw counts of the top hit dataset

These plots show the changes in the gene count of the top hits w.r.t to the timepoint. Our results showcase that all the genes have consistent enrichment in the E2-YPD condition as we increase the time of treatment
```{r}
#visualization
#ERG11

p1 <- ggplot(Metadata, aes(x = Timepoints, y = ERG11, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "ERG11 counts Variation",
       x = "Experiment", y = "ERG11 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10()+
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#FLR1
p2 <- ggplot(Metadata, aes(x = Timepoints, y = FLR1, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "FLR1 counts Variation",
       x = "Experiment", y = "FLR1 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10()+
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#PDR5
p3 <- ggplot(Metadata, aes(x = Timepoints, y = PDR5, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "PDR5 counts Variation",
       x = "Experiment", y = "PDR5 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10()+
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#YAP1
p4 <- ggplot(Metadata, aes(x = Timepoints, y = YAP1, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "YAP1 counts Variation",
       x = "Experiment", y = "YAP1 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10()+
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#MGE1
p5 <- ggplot(Metadata, aes(x = Timepoints, y = MGE1, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "MGE1 counts Variation",
       x = "Experiment", y = "MGE1 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10()+
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#pdf("filename1.pdf", width = 16, height = 16)
gridExtra::grid.arrange(p1, p2, p3, p4, p5, nrow =2)
#dev.off()
```
Sensitivity
```{r}
#AUS1
p6 <- ggplot(Metadata, aes(x = Timepoints, y = AUS1, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "AUS1 counts Variation",
       x = "Experiment", y = "AUS1 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10() +
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#VMA16
p7 <- ggplot(Metadata, aes(x = Timepoints, y = FRE3, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "FRE3 counts Variation",
       x = "Experiment", y = "FRE3 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10() +
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

p8 <- ggplot(Metadata, aes(x = Timepoints, y = HSP104, color = Conditions, group = Conditions)) +
  geom_jitter() +
  labs(title = "HSP104 counts Variation",
       x = "Experiment", y = "HSP104 counts") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text( size = 18),
        axis.title = element_text(size = 20),
        plot.title = element_text(size = 20),
        legend.title = element_text(size=20),
        legend.text = element_text(size=16))+
  scale_y_log10() +
  geom_smooth(method = "lm", alpha = .15, aes(fill = Conditions))

#pdf("filename.pdf", width = 8, height = 12)
gridExtra::grid.arrange(p6,p7,p8)
#dev.off() 

```
Added by Saaz 6/24/26

```{r visualisationforpaper}
library(readxl)
library(rapportools)
library(openxlsx)
library(forcats)
library(ggplot2)
library(gghighlight)
library(ggrepel)

Gene_annotation <- read_excel("GCF_000146045.2_R64_annotation.xlsx")
Sig_hits=decreasing_order_of_significant_genes

#Annotate - annotate each hit gene with the Chromosome it is on
Sig_hits$Locus_tag=rownames(Sig_hits)

l=length(Sig_hits$gene)

relchr=vector(length=l)

for (i in (1:l)) {
  relname=Sig_hits$Locus_tag[i]
  relrow=which(Gene_annotation$Locus_tag==relname)
  if (is.empty(relrow)) {
    relchr[i]=FALSE
  } else {
    relchr[i]=Gene_annotation$Chromosome[relrow]
  }
}

Sig_hits$Chr=relchr

#Showing the nunber of sig hits on each chr
table(Sig_hits$Chr)

#Plot 72h vs 48h hits, with only Chr 8 hits coloured
ggplot(Sig_hits,aes(x=time_48h,y=time_72h))+geom_point(aes(color=Chr))+gghighlight(Chr=="VIII")+theme_bw()

#Plot 72h results alone for all chromosomes separately
Sig_hits_focus=Sig_hits
Sig_hits_focus$Chrfilter="Other Chr"
Sig_hits_focus$Chrfilter[which(Sig_hits_focus$Chr=="VIII",)]="Chr VIII"
Sig_hits_focus$Genefilter=Sig_hits_focus$Chrfilter
Sig_hits_focus$Genefilter[1:5]="Sighit"

Sig_hits_focus$Chrnumber=as.roman(Sig_hits_focus$Chr)
Sig_hits_focus$Chrnumber=as.numeric(Sig_hits_focus$Chrnumber)
Sig_hits_focus=Sig_hits_focus[!is.na(Sig_hits_focus$Chrnumber),]
Sig_hits_focus=Sig_hits_focus[abs(Sig_hits_focus$time_72h)>1.5,]

Sig_hits_bychr=ggplot(Sig_hits_focus,aes(x=fct_inseq(as.factor((Chrnumber))),y=time_72h,color=Chr)) + geom_jitter(width=0.1,show.legend=FALSE) + theme_bw() + ylab("log2 abundance (YPD+Fluconazole / YPD)") + xlab("Chromosome") + theme(legend.title=element_blank(),text=element_text(size=13,face="bold")) + guides(color="none", size="none") + geom_text_repel(aes(label=ifelse(Genefilter=="Sighit",as.character(gene),'')),hjust=-0.2,vjust=0.1)

Sig_hits_bychr

ggsave("./Sig_hits_bychr.pdf", plot = Sig_hits_bychr, device = "pdf")
ggsave("./Sig_hits_bychr.tiff", plot = Sig_hits_bychr, device = "tiff")

```
