---
title: "Growth curve analysis using the Tecan 96-well plate reader"
author: "Saaz Sakrikar, modifying code from Grace Avecilla and David Gresham"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
---

The experiment below includes: 2 strains of S. cerevisiae, grown in 6 different concs of FLC (including no FLC). The strains are DGY3612 (MATa/alpha mCherry-HygB near ERG11/mCitrine-KanMX near ERG11) and DGY3 (MATa/alpha). I am gathering data from diploid ancestral/naive strain to obtain a fluconazole IC50 for diploid S. cerevisiae.

The data comprise 288 OD measurements that were acquired over a 24 hour period.

# Growth curves requirements


* OD data from the Tecan in a csv with each well's data in a column, the first row contained the well names, and the first column is time in seconds 

* A sample sheet, with the four required columns: __Well__, __Sample__,	__Bio_rep__,	__Tech_rep__
  + The Well column did not have 0s in the well names (i.e. A1, not A01)
  + For blank wells, we put 'B' in the Sample column of the sample sheet
  + For blank wells, we put 'NA' for all other column values
  + Additional columns are possible. In this case, "Strain", "Type" (2-copy or 1-copy (in relation to ERG11)), "FLC_conc".



```{r input, warning=FALSE,message=FALSE}
library(tidyverse)
library(growthcurver)

#theme_set(theme_minimal())

cycles = 288
samplesheet = read_csv("Apr29_meta.csv")
#knitr::kable(samplesheet)

data_od = read_csv("Apr29_growth.csv",skip = 0, n_max = cycles) 
colnames(data_od)[1] <- "time"
#convert time to hrs from seconds
data_od$time <- as.numeric(data_od$time)/60/60

#knitr::kable(data_od)


# output csv name, used at the end only.
out = 'output.csv'

```

Transformation of data for plotting

```{r setup}
data_od_for_plot <- data_od %>% 
  pivot_longer(cols = -time, names_to = "well", values_to = "od") %>% 
  left_join(samplesheet, by = c("well" = "Well")) %>%
  unite(Sample_Well, well, Sample, sep = " ", remove = FALSE)

```

## Individual Growth Profiles

We  plot the entire dataset arrayed in 96 well format it appears on the Tecan for visual inspection.  This allows us to easily identify any wells that should be excluded.

```{r all_well_plot}
data_od_for_plot %>%
  mutate(Sample_Well = str_replace(Sample_Well, "(?<![0-9])([0-9])(?![0-9])", "0\\1")) %>%
  ggplot(aes(time, od, color = as.factor(Type))) +
  geom_line() +
  facet_wrap(~Sample_Well, ncol=12) +
  xlab("Time (hrs)") + scale_y_continuous(breaks=c(0, 2)) + scale_x_continuous(breaks=c(0, 12, 24)) + theme_bw()

```

#Obtain growth paramaters using logisitc curve

```{r analysis}
gc_out <- SummarizeGrowthByPlate(data_od, 
                                 plot_fit = FALSE) %>% 
  left_join(samplesheet, by = c("sample" = "Well")) %>%
  rename(well = sample, sample = Sample)

#knitr::kable(gc_out)

gc_sampleout=gc_out[!is.na(gc_out$Strain),]
write.csv(gc_sampleout,paste(out))
```

Plot the growth rate r and area under the curve AUC

```{r plotting}

ggplot(gc_sampleout,aes(x=Strain,y=r)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=Strain, color=Strain)) + facet_wrap(~FLC_conc) + theme_bw() + ggtitle("Growth rate r at different FLC concs")

#ggplot(gc_sampleout,aes(x=Type,y=t_gen)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=Strain, color=Strain)) + facet_wrap(~FLC_conc) + theme_bw() + ggtitle("Doubling time tgen at different FLC concs")

ggplot(gc_sampleout,aes(x=Strain,y=auc_l)) + geom_dotplot(binaxis="y", stackdir="center", dotsize=2, aes(fill=Strain, color=Strain)) + facet_wrap(~FLC_conc) + theme_bw() + ggtitle("Area under the curve AUC at diff FLC concs")

ggplot(gc_sampleout,aes(x=as.factor(FLC_conc),y=r)) + geom_dotplot(binaxis="y", stackdir="center", aes(fill=Strain, color=Strain)) + facet_wrap(~Type) + theme_bw() + ggtitle("Growth rate r for diff strain types")
```
